Concurrent Programming with CXL

### 2020-06-06 book/paper.tex
Working on producer/consumer

A good example is the so-called \emph{bounded buffer}.  It is essentially
a queue implemented using a circular buffer of a certain length and two pointers:
one where items are inserted and one from where items are extracted.  If the
buffer is full, processes that want to add more items have to wait.  In the usual
formulation, processes that want to extract an item when the buffer is empty
also have to wait.
This problem is known as the ``Producer/Consumer Problem.''
It would be easy to solve using a busy waiting
approach, but again, we would like to avoid busy waiting.

Introduced by the famous Dutch computer scientist Edsger Dijkstra,
a \emph{semaphore} is a primitive that fits the bill.  A semaphore is essentially
a counter that can be incremented and decremented but is not allowed to go
below zero.  The semaphore counter is typically initialized to the number of
units of a resource initially available.
When allocating a resource, a process decrements the
counter using the \texttt{procure} or simply \texttt{P} operation.  The
\texttt{procure} operation blocks the invoking process if the counter is zero.
To release the resource, a process increments the resource using the
\texttt{vacate} or {V} operation.
Figure~\ref{fig:semaphore} shows the \texttt{synch} module implementation of
semaphores.

@@

>>>>

A good example is the so-called \emph{bounded buffer}.  It is essentially
a queue implemented using a circular buffer of a certain length and two pointers:
one where items are inserted and one from where items are extracted.  If the
buffer is full, processes that want to add more items have to wait.  In the usual
formulation, processes that want to extract an item when the buffer is empty
also have to wait.
This problem is known as the ``Producer/Consumer Problem.''

Figure~\ref{fig:pc} gives a high-level description of what we want.  The
two numbers in a state specify
the number of items that the producers still can produce and
the number of items that the consumers have consumed.  In this case
there are three items to produce initially.

It would be easy to solve using a busy waiting
approach, but again, we would like to avoid busy waiting.
Introduced by the famous Dutch computer scientist Edsger Dijkstra,
a \emph{semaphore} is a primitive that fits the bill.  A semaphore is essentially
a counter that can be incremented and decremented but is not allowed to go
below zero.  The semaphore counter is typically initialized to the number of
units of a resource initially available.
When allocating a resource, a process decrements the
counter using the \texttt{procure} or simply \texttt{P} operation.  The
\texttt{procure} operation blocks the invoking process if the counter is zero.
To release the resource, a process increments the resource using the
\texttt{vacate} or {V} operation.
Figure~\ref{fig:semaphore} shows the \texttt{synch} module implementation of
semaphores.

@@

We can run the code through CXL, but what does it test?  There are no
assert statements in the code.
We can split the functionality of this code into two pieces.  One is
the synchronization part: producers should never run more than \texttt{NSLOTS}
ahead of the consumers, but they should be able to produce up to \texttt{NSLOTS}
items even if there are no consumers.  The other part is the data: we want
to make sure that the items are sent through the buffer in order and without
loss.  We will first focus on the synchronization.

[RVR.  More here.]
